---
title: "Simal Data Package Assignment"
author: "Simal Sami"
format: html
editor: visual
---

# Load Packages

```{r}
#| message: false
#| echo: false
#| include: false
library(tidyverse)
library(ggthemes)
library(readr)
library(knitr) 
library(kableExtra)

theme_set(theme_minimal(base_size = 12))

```

# Load Data

```{r}
#| message: false
movies <- read_csv("Data/movies.csv")
raw_bechdel <- read_csv("Data/raw_bechdel.csv")

```

```{r}
#| message: false
source("R/get_movies_data.R")
```

# Section 1: Data Sources

**Background**

The Bechdel test is a test used to test gender biases and representation of women in movies, TV shows, etc. Often women in film portray characters that lack depth or are depicted as certain archetypes of damsels in distress. The test was first introduced by cartoonist Alison Bechdel who said that in order to pass the test a work of fiction must:

-   Involve two named women characters

-   The two women must converse with one another

-   The conversation does not involve a man or any mention of a male character.

Many movies fail to pass this test and it highlights the need for more female representation in strong roles beyond male-centric themes.

**The Data**

The data is from a public data source call [TidyTuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-09/readme.md) and it was created by [FiveThirtyEight](https://github.com/fivethirtyeight/data/tree/master/bechdel), a polling aggregation website using [Bechdel](https://bechdeltest.com/) test data which analyzed over 5,000 films and whether or not they passed the test, alongside [TheNumbers.com](https://www.the-numbers.com/) which compiled a data set of various stats on the different movies including awards, directors, ratings, revenue, etc. The data was collected to expose gender biases within film and compare whether or not those films that pass or do not pass the Bechdel test differ in budgets, ratings, revenue, etc. The set analyzes 1,794 films between the years of 1970-2013. The films in the data set are an aggregate of the intersection between the films in the Numbers data set and the Bechdel data set.

# Format of Data

The data is tabular and stored as a csv file.

# Data Structure

## Raw Bechdel Data:

The Raw Bechdel data set contains 10,136 observations in tabular form. Each row of the Raw Bechdel data set is a movie which contains the following 5 columns:

-   A Movie ID

-   The year it was made

-   The IMDB ID which is an ID assigned to each movie by the Internet Movie Database

-   The title of the movie

-   A Rating which is a scale of 0-3:

    -   1 means the movies features at least 2 named women.

    -   2 means that the women converse with one another

    -   3 means that the conversation is about something other than a man.

    -   0 means the movie is un-scored or contains none of the Bechdel features.

Below is a preview of the dataset:

```{r}
#| message: false
#| label: tbl-Raw-Bechdel
#| tbl-cap: Preview of Raw Bechdel Dataset

raw_bechdel %>%
  head(5) %>%
  kable()
```

## Movies Data:

The Movies data set contains 1,794 observations. The data set is the the product of an intersection between the raw bechdel data set (every movie that has been put through the Bechdel test by the Bechdel website) and the data set from the [Numbers](https://www.the-numbers.com/) website which is a congregated data set of every movie to hit the box office and their various characteristics.

```{r}
#| message: false
#| label: tbl-movies
#| tbl-cap: Preview of Movies Data set

movies %>%
  head(5) %>%
  kable()
```

# Data Issues

The data does not contain any errors or missing values.

# Data Cleaning

The data has been cleaned using a script provided by TidyTuesday it can be found it the repository under the R folder. The script does the following:

-   Fetches the Bechdel data

-   Converts the JSON data into a tidy data frame and filters the movies by years.

-   Writes this data to a CSV file called raw_bechdel

-   Fetches the movie data.

-   Converts the JSON data into a tidy data frame and also converts columns into their proper data type.

-   Writes the data frame to a CSV called combo_movies

-   Joins together the combo_movies data set and raw-bechdel data set based on the movies they have in common and creates a new data frame written to movies.csv
